### 2024-03-29 paper/info_sheet.tex
added info sheet

\title{MLVU information sheet}
\author{}
\date{}
\maketitle

\noindent \textit{Please include this page in your report either at the start or at the end, before the appendix. Do not change the formatting.}

\paragraph{Group number}~

02

\paragraph{Authors}~

\begin{tabular}{r r}
name &student number \\
\hline
Kai Niermann & 2720905 \\
Dávid Miklo & 2767001 \\
Trix Taiclet & 2785891 \\
Conner Dassen & 2695049 \\
Red Kaláb & 2735630 \\
\end{tabular}

\paragraph{Software used} \textit{jupyter, python3.10, keras3, tf2.16, various data viz libraries, gcp compute engine to train the models, pandas for most of the data science related tasks}

\paragraph{Use of AI tools} \textit{}
\paragraph{Link to code (optional)} \textit{}

\paragraph{Group disagreements} \textit{}